---
title: "Anonymous functions in dplyr and purrr"
author: "Hao Ye"
date: "`r format(Sys.Date(), '%Y %B %d')`"
output: 
  github_document:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preamble

```{r load packages}
library(dplyr)
library(purrr)
```

# Simple Workflow

This is a silly example, but demonstrates what we want to do:

```{r single dataset workflow}
sample_data <- data.frame(x = runif(10), y = runif(10))
print(sample_data)

# imagine F is some model that we're fitting that returns some complex object, 
# in this case, computing a prediction for y using a linear model
F <- function(df) {
    data.frame(y_obs = df$y, 
               y_hat = lm(y ~ x, data = df)$fitted.values)
}

sample_output <- F(sample_data)
print(sample_output)

# imagine we want to process the output by computing the residual
G <- function(output) {
    data.frame(output, 
               resid = output$y_obs - output$y_hat)
}

sample_summary <- G(sample_output)
print(sample_summary)
```

# Fancier Workflow

Suppose that we have multiple datasets that we plan to do the same procedure to:

```{r multiple dataset workflow}
data_list <- list(data.frame(x = runif(10), y = runif(10)), 
                     data.frame(x = runif(100), y = runif(100)), 
                     data.frame(x = runif(1000), y = runif(1000)))

output_list <- lapply(data_list, F)
summary_list <- lapply(output_list, G)
```

# Workflow with dplyr and purrr

We can replace the `lapply` calls using the `map` function from the `purrr` package:

```{r multiple dataset workflow using purrr}
# note that we use the data_frame instead of data.frame:
# the former makes it simpler to have a list column and makes a tibble object
my_data <- data_frame(species = c("A", "B", "C"), 
                      data = data_list)
print(my_data)

results <- my_data %>% 
    mutate(output = map(data, F)) %>% 
    mutate(summary = map(output, G))

print(results)
```

# Using a dplyr workflow to process model output

## single dataset example

Note that our summary function `G` can be rewritten simply using `mutate`.

```{r}
sample_summary_2 <- sample_output %>% 
    mutate(resid = y_obs - y_hat)

print(sample_summary_2)
```

## multiple dataset example

But if we want to use it with map, we need to make an anonymous function. The key here is to wrap the `dplyr` workflow in parantheses:

```{r}
G_dplyr <- function(output) {
    (output %>% mutate(resid = y_obs - y_hat))
}

results_dplyr <- my_data %>% 
    mutate(output = map(data, F)) %>%
    mutate(summary = map(output, G_dplyr))

print(results_dplyr)
```

Yes, it's silly here, becasue `G_dplyr` is no more complex than `G`, but if we have some complex workflow in `dplyr` that we need to map, this allows us to do that.

