---
title: "D-stability of interaction matrices"
author: "Hao Ye"
date: "`r format(Sys.Date(), '%Y %B %d')`"
output: 
  github_document
---

## Definitions

Suppose we have an *interaction* matrix, $A$ that is defined as:
$$A_{ij} = \left.\frac{\partial\left(\frac{1}{N_i}\frac{dN_i}{dt}\right)}{\partial N_j}\right|_{N^*}$$

evaluated at some equilibrium point, $N^*$.

Then, the corresponding *community* matrix, $C$ is:
$$C_{ij} = \left.\frac{\partial\left(\frac{dN_i}{dt}\right)}{\partial N_j}\right|_{N^*} = {N_i}^* A_{ij}$$

Furthermore, a matrix, $M$, is *stable* (asymptotic stability, Lyapunov stability) if the real part of each eigenvalue is negative:
$$\forall \lambda_M : \operatorname{Re}{\lambda} < 0$$

and a matrix, $M$, is *D-stable* if $DM$ is stable for any positive diagonal $D$ (all entries on the main diagonal are positive, and the rest are 0). Note that if $M$ is D-stable, it is also stable (set $D$ to the identity matrix.)

## Example

Not all stable matrices are D-stable. From Logofet (2005), we borrow the example of a matrix with a particular sign pattern:

```{r}
set.seed(45)
A <- matrix(c(abs(rnorm(1)), -abs(rnorm(1)), 
              abs(rnorm(1)), -abs(rnorm(1))), byrow = TRUE, nrow = 2)
cat(A)
```

Is `A` stable?
```{r}
lambda <- eigen(A)$values
cat(lambda)
all(Re(lambda) < 0)
```

What if we multiply `A` by a positive diagonal matrix?
```{r}
N <- diag(runif(2) * c(100, 10) + c(100, 10))
C <- A %*% N
```

Is `C` stable?

```{r}
lambda <- eigen(C)$values
cat(lambda)
all(Re(lambda) < 0)
```