---
title: "s-map coefficients check"
author: "Hao Ye"
date: "`r format(Sys.Date(), '%Y %B %d')`"
output: 
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##

First, create an S-map model using some sample data:

```{r}
library(rEDM)
data("block_3sp")

lib <- c(1, 100)
pred <- c(101, 200)
columns <- c("x_t-1", "x_t-2")

smap_out <- block_lnlp(block_3sp, lib = lib, pred = pred, 
                       method = "s-map", tp = 0, theta = 1, 
                       columns = columns, target_column = "x_t", 
                       stats_only = FALSE, save_smap_coefficients = TRUE, 
                       short_output = TRUE)
```

##

Next, pull out the s-map coefficients and input variables:

```{r}
predictions <- smap_out[[1]]$model_output$pred
smap_coeffs <- smap_out[[1]]$smap_coefficients

input_block <- block_3sp[pred[1]:pred[2], columns]
```

##

Now, try and reproduce the predictions by multiplying the input block with the s-map coefficients. Note that because there is a constant term, we add a constant 1 to the right side of the input block (in prep. for the eventual matrix product).

```{r}
# each row of inputs to be multipled by corresponding s-map coefficints
schur_product <- cbind(input_block, 1) * smap_coeffs

# compute sum of rows
calculated_preds <- rowSums(schur_product)
```

##

The calculations have a names attribute that is not in the extracted predictions, so we clear it here for the equality test.

```{r}
names(calculated_preds) <- NULL

# Test for equality
message("predictions == calculated_preds?  ", 
        all.equal(predictions, calculated_preds))
```




