---
title: "Vaccination Rate by Zodiac Sign"
author: "Hao Ye"
date: "10/20/2021"
output: github_document
---

```{r setup, message = FALSE}
library(tidyverse)
library(lubridate)
```

# SaltLakeHealth tweet

https://twitter.com/SaltLakeHealth/status/1450570095550517249

```{r}
SLC_vacc_prop <- data.frame(zs = c("Leo", "Aquarius", 
                                 "Aries","Sagittarius", 
                                 "Cancer","Taurus", 
                                 "Gemini","Libra", 
                                 "Pisces","Capricorn", 
                                 "Virgo","Scorpio"), 
                          vacc_rate = c(0.70, 0.67, 
                                   0.59, 0.59, 
                                   0.58, 0.56, 
                                   0.55, 0.54, 
                                   0.51, 0.51, 
                                   0.50, 0.46))

# check that we got the input correct
knitr::kable(SLC_vacc_prop)
```

# Back-calculate proportion vaccinated

We assume the calculations are as follows:

```{r}
# SLC_vacc_prop$vacc_rate = num_vac / (zs_prop * n)
#   - num_vac is from vaccination data
#   - zs_prop is zodiac sign proportion (from some source)
#   - n is SLC population estimate
```

## zodiac sign proportion

The tweet suggests the zodiac sign proportion is from University of Texas-Austin. The reddit thread at https://www.reddit.com/r/SaltLakeCity/comments/qbzwra/helpful_data_from_salt_lake_county_health/ points to this data from Texas A&M:
http://faculty.tamucc.edu/sfriday/wordpress/?p=1317#:~:text=Most%20Common%20to%20Least%20Common%20Zodiac%20Signs%20,%200.090%20%25%20%208%20more%20rows%20

```{r}
tamu_zs_prop <- data.frame(zs = c("Scorpio", "Virgo", 
                                  "Gemini","Pisces", 
                                  "Libra","Cancer", 
                                  "Taurus","Capricorn", 
                                  "Aries","Sagittarius", 
                                  "Leo","Aquarius"), 
                           zs_prop = c(0.094, 0.093, 
                                    0.092, 0.090, 
                                    0.087, 0.084, 
                                    0.083, 0.082, 
                                    0.081, 0.073, 
                                    0.071, 0.063))

# check that we got the input correct
knitr::kable(tamu_zs_prop)
```

## Back-calculation

```{r}
SLC_vacc_df <- SLC_vacc_prop %>%
    left_join(tamu_zs_prop, by = "zs") %>%
    mutate(num_vac_over_n = vacc_rate * zs_prop)

knitr::kable(SLC_vacc_df)
```

# Better zodiac sign proportion data

FiveThirtyEight has US birth data at https://github.com/fivethirtyeight/data/tree/master/births

Let's check the distribution of zodiac signs:

## Function to return zodiac sign from date input

Originally from https://stackoverflow.com/questions/48026462/writing-r-function-using-lubridate-package-that-takes-dateex-august-14-and, with some modifications to vectorize properly
```{r}
zodiac_sign <- function(input)
{
    zodiac_sign_df <- data.frame(Month = c("March", "April", "May", "June", "July", "August", "September", "October", "November", "December", "January", "February"),
                                 Zodiac_Separation_Dt = c(21, 20, 21, 21, 23, 23, 23, 23, 22, 22, 20, 19),
                                 LowerSideZS = c("Pisces", "Aries","Taurus","Gemini","Cancer","Leo","Virgo","Libra","Scorpio","Sagittarius","Capricorn","Aquarius"),
                                 UpperSideZS = c("Aries","Taurus","Gemini","Cancer","Leo","Virgo","Libra","Scorpio","Sagittarius","Capricorn","Aquarius", "Pisces"),
                                 stringsAsFactors = FALSE )
    val_df <- zodiac_sign_df[match(months(input), zodiac_sign_df$Month), ]
    ifelse( day(input) >= val_df$Zodiac_Separation_Dt, 
            val_df$UpperSideZS, val_df$LowerSideZS)
}
```

## FiveThirtyEight US birth data

There are two datasets, let's check both:
```{r, results = "hold"}
SSA_data <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/births/US_births_2000-2014_SSA.csv"
CDC_NCHS_data <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/births/US_births_1994-2003_CDC_NCHS.csv"

SSA_dat <- read.csv(SSA_data) %>%
    mutate(date = ISOdate(year, month, date_of_month), 
           zs = zodiac_sign(date)) %>%
    count(zs, wt = births) %>%
    mutate(zs_prop = n / sum(n)) %>%
    arrange(desc(zs_prop))

CDC_NCHS_dat <- read.csv(CDC_NCHS_data) %>%
    mutate(date = ISOdate(year, month, date_of_month), 
           zs = zodiac_sign(date)) %>%
    count(zs, wt = births) %>%
    mutate(zs_prop = n / sum(n)) %>%
    arrange(desc(zs_prop))

knitr::kable(SSA_dat)
knitr::kable(CDC_NCHS_dat)
```

These both have very similar zs_prop values.

# Re-calculate  proportion vaccinated

Using `SSA_dat$zs_prop`:
```{r}
SLC_vacc_df %>%
    select(zs, num_vac_over_n) %>%
    left_join(SSA_dat, by = "zs") %>%
    mutate(vacc_rate = num_vac_over_n / zs_prop) %>%
    select(zs, vacc_rate) %>%
    arrange(desc(vacc_rate)) %>%
    knitr::kable()
```

Using `CDC_NCHS_dat$zs_prop`:
```{r}
SLC_vacc_df %>%
    select(zs, num_vac_over_n) %>%
    left_join(CDC_NCHS_dat, by = "zs") %>%
    mutate(vacc_rate = num_vac_over_n / zs_prop) %>%
    select(zs, vacc_rate) %>%
    arrange(desc(vacc_rate)) %>%
    knitr::kable()
```

Ok, well, both of those actually look much more reasonable in terms of disparities in vaccination rate with zodiac sign.
